Draft_figures_2020-06-02

### 2020-06-04 R/Draft_figures_2020-06-02.Rmd
New back to back plots of flowering in shortening vs lengthening days

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
```

@@

daylength_change <- fl50_dyln %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
                               MEAS <= 172 ~ "Lengthening",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP, PLANT_ID) %>%
  summarise(count = n()) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP) %>%
  tally() #%>%
  #group_by(SITE, SUBPOP) %>%
  #mutate(frac = n/sum(n)) %>%
  #pivot_wider(names_from = SUBPOP, values_from = n) %>%
  #left_join(sites) %>%
dyln_change <- daylength_change %>%
  ggplot(aes(x = FL_B4_SOL, y = n, width = 0.7)) + 
  geom_bar(aes(fill = SUBPOP), stat = "identity", position = position_dodge()) +
  facet_wrap(~SITE, nrow = 2) +
  scale_fill_viridis(discrete = TRUE) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "right", panel.spacing.x = unit(x = 0.01, unit = "npc")) +
  ylab("Count of Genotypes") +
  xlab("Does the genotype flower when days are 
       lengthening or shortening?") 
save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = dyln_change)

?unit
?geom_bar
```
At Kingsville, all five subpopulations had individuals that flowered while days were lengthening. Only the admixed 4X and the Gulf subpops had a majority of individuals that flowered while days were shortening (55% and 85.3%).
At Pickle, all five subpopulations had individuals that flowered while days were lengthening. Only the Gulf subpop had a majority of individuals that flowered while days were shortening (94.0%).
At Temple,  all five subpopulations had individuals that flowered while days were lengthening. The admixed 4X, Atlantic, and Gulf had a majority of individuals that flowered while days were shortening (65.2%, 57.9%, and 99.5%).
At Stillwater, 4X, Atlantic, and Midwest individuals flowered while days were lengthening, but these individuals were in the minority (5.3%, 2.3%, 3.7%).
At Columbia, 8X and Midwest individuals flowered while days were lengthening (1.4% and 5%).
Further north than Columbia, virtually all individuals flowered while days were shortening.

```{r}
fl50_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

```

>>>>

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE)) 
phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))

subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))

# gulf: #F47F72, atlantic: #6E91CB, midwest: #442C83
```

@@

daylength_change <- fl50_dyln %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
                               MEAS <= 172 ~ "Lengthening",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP, PLANT_ID) %>%
  summarise(count = n()) %>%
  group_by(SITE, FL_B4_SOL, SUBPOP) %>%
  tally() #%>%
  #group_by(SITE, SUBPOP) %>%
  #mutate(frac = n/sum(n)) %>%
  #pivot_wider(names_from = SUBPOP, values_from = n) %>%
  #left_join(sites) %>%
daylength_change$SUBPOP <- factor(daylength_change$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
color_code <- c("grey", "#6E91CB","#F47F72", "#442C83", "#090906")
dyln_change <- daylength_change %>%
  left_join(sites) %>%
  left_join(subpops) %>%
  dplyr::select(SITE, FL_B4_SOL, SUBPOP, n) %>%
  ungroup() %>%
  #add_row(SITE = "BRKG", FL_B4_SOL = "Shortening") %>%
  #add_row(SITE = "LINC", FL_B4_SOL = "Shortening") %>%
  complete(SITE, FL_B4_SOL, SUBPOP, fill = list(n = 0))
delta_dyln %>%
  ggplot(aes(x = FL_B4_SOL, y = n, width = 0.7)) + 
  geom_bar(aes(fill = SUBPOP), stat = "identity", position = position_dodge()) +
  facet_wrap(~manu_site, nrow = 2) +
  scale_fill_manual(values = color_code) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "right", panel.spacing.x = unit(x = 0.01, unit = "npc")) +
  ylab("Count of Genotypes") +
  xlab("Does the genotype flower when days are 
       lengthening or shortening?") 
save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = delta_dyln)

```
```{r}
dyln_change$SUBPOP <- factor(dyln_change$SUBPOP, levels = rev(c("4X", "Atlantic", "Gulf", "Midwest", "8X")))
color_code <- rev(c("grey", "#6E91CB","#F47F72", "#442C83", "#090906"))

site_subpop_list <- list()
eightsites <- c("KING", "PKLE", "TMPL", "STIL", "CLMB", "LINC", "KBSM", "BRKG")

for(i in seq_along(eightsites)){
p1 <- dyln_change %>%
  filter((SITE %in% eightsites[i]) & FL_B4_SOL == "Lengthening") %>%
  ggplot(aes(x = SUBPOP, y = n, width = 1)) + 
  geom_col(aes(fill = SUBPOP), position = "dodge") +
  scale_y_reverse(expand = c(0.002, 0.002), limits = c(240, 0)) + coord_flip() +
  scale_fill_manual(values = color_code) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm"),
        plot.margin = unit(c(5.5, 0, 5.5, 5.5), "pt")) +
  ylab("Genotypes") +
  xlab("Subpopulation") + labs(subtitle = "Lengthening", title = sites$manu_site[sites$SITE == eightsites[i]])
p2 <- dyln_change %>%
  filter((SITE %in% eightsites[i]) & FL_B4_SOL == "Shortening") %>%
  ggplot(aes(x = SUBPOP, y = n, width = 1)) + 
  geom_col(aes(fill = SUBPOP), position = "dodge") +
  scale_y_continuous(expand = c(0.002, 0.002), limits = c(0,240)) +
  coord_flip() +
  scale_fill_manual(values = color_code) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        axis.text.y = element_blank(), axis.title.y = element_blank(),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm"),        axis.line.y = element_blank(), axis.ticks.y = element_blank(),
        plot.margin = unit(c(5.5, 5.5, 5.5, -3.5), "pt")) +
  geom_hline(yintercept = 0, linetype = 2) + labs(subtitle = "Shortening", title = " ") + ylab(" ")

site_subpop_list[[i]] <- plot_grid(p1, p2, rel_widths = c(1.66,1))
}

subpop_dyln_plot <- plot_grid(site_subpop_list[[1]], site_subpop_list[[2]], site_subpop_list[[3]],
          site_subpop_list[[4]], site_subpop_list[[5]], site_subpop_list[[6]],
          site_subpop_list[[7]], site_subpop_list[[8]], nrow = 2)

save_plot(filename = "../manuscript/plots/daylength_change_subpops_back_to_back.png", plot = subpop_dyln_plot, base_asp = 2.2, base_height = 5)
```

At Kingsville, all five subpopulations had individuals that flowered while days were lengthening. Only the admixed 4X and the Gulf subpops had a majority of individuals that flowered while days were shortening (55% and 85.3%).
At Pickle, all five subpopulations had individuals that flowered while days were lengthening. Only the Gulf subpop had a majority of individuals that flowered while days were shortening (94.0%).
At Temple,  all five subpopulations had individuals that flowered while days were lengthening. The admixed 4X, Atlantic, and Gulf had a majority of individuals that flowered while days were shortening (65.2%, 57.9%, and 99.5%).
At Stillwater, 4X, Atlantic, and Midwest individuals flowered while days were lengthening, but these individuals were in the minority (5.3%, 2.3%, 3.7%).
At Columbia, 8X and Midwest individuals flowered while days were lengthening (1.4% and 5%).
Further north than Columbia, virtually all individuals flowered while days were shortening.

Latitude of origin bins for flowering while day lengthening/shortening plot

```{r}
fl50_dyln %>%
  filter(!is.na(LATITUDE)) %>%
  mutate(latbin = floor(abs(LATITUDE)/5)) %>%
  filter(WHERE == "GWAS") %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
                               MEAS <= 172 ~ "Lengthening",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, latbin) %>% tally()

dyln_latbin <- fl50_dyln %>%
  filter(!is.na(LATITUDE)) %>%
  mutate(latbin = floor(abs(LATITUDE))) %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "Shortening",
                               MEAS <= 172 ~ "Lengthening",
                               TRUE ~ NA_character_)) %>%
  group_by(SITE, FL_B4_SOL, latbin, PLANT_ID, SUBPOP) %>%
  summarise(count = n()) %>%
  dplyr::select(-PLANT_ID) %>%
  complete(SITE, FL_B4_SOL, latbin, SUBPOP) %>%
  group_by(SITE, FL_B4_SOL, latbin, SUBPOP) %>%
  summarise(count = sum(!is.na(count))) %>%
  complete(SITE, FL_B4_SOL, latbin, SUBPOP, fill = list(count = 0)) %>%
  arrange(latbin)

#dyln_change <- 
dyln_latbin %>%
  left_join(sites, by = "SITE") %>%
  filter(!(SITE %in% c("FRMI", "OVTN"))) %>%
  ggplot(aes(x = FL_B4_SOL, y = count, width = 0.8)) + 
  geom_bar(aes(fill = as_factor(latbin)), stat = "identity", position = position_dodge()) +
  facet_wrap(~manu_site, nrow = 2) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "right", panel.spacing.x = unit(x = 0.01, unit = "npc")) +
  ylab("Count of Genotypes") +
  xlab("Do genotypes from that latitude of origin flower when days are 
       lengthening or shortening?") 
save_plot(filename = "../manuscript/plots/daylength_change_subpops.png", plot = dyln_change)
dyln_latbin2 <- dyln_latbin %>% ungroup() %>% add_row(latbin = 24, count = 0) %>%
  complete(SITE, FL_B4_SOL, latbin, SUBPOP, fill = list(count = 0)) %>%
  filter(!is.na(FL_B4_SOL) & !is.na(SUBPOP)) %>%
  mutate(count = case_when(latbin == 24 ~ 48,
                           TRUE ~ count)) %>%
  unique()
```

```{r}
latranges <- metadata %>%
  group_by(SUBPOP) %>%
  summarise(min = min(LATITUDE, na.rm = TRUE),
            max = max(LATITUDE, na.rm = TRUE))

metadata$SUBPOP <- factor(metadata$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
paneld <- metadata %>%
  ggplot(aes(x = SUBPOP, y = LATITUDE)) +
  geom_violin(aes(color = SUBPOP)) + geom_jitter(aes(color = SUBPOP), alpha = 0.1, height = 0, width = 0.1) + scale_color_manual(values = rev(color_code)) + theme(legend.position = "none", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) + ylab("Latitude of origin") + xlab("Subpopulation") + labs(title = "", subtitle = "")
```

```{r}

panela <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("KING"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX1", subtitle = "Day length change on date of flowering") +
  xlim(c(25,45))
panelb <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("PKLE"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX2", subtitle = "Day length change on date of flowering") +
  xlim(c(25,45))
panelc <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("TMPL"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = as_factor(latbin)), position = "dodge") +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX3", subtitle = "Day length change on date of flowering") +
  xlim(c(25,45))

dyln_lat_plot <- plot_grid(paneld, panela, panelb, panelc, nrow = 1)
save_plot(filename = "../manuscript/plots/Daylength_changes_flowering_date_by_latitude_of_origin_subpops.svg", plot = dyln_lat_plot, base_asp = 2.6, base_height = 4)
```

```{r}
filter(unique(dyln_latbin2), count > 200)
panela <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("KING"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = SUBPOP)) +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_manual(values = rev(color_code)) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX1", subtitle = "Day length change on date of flowering") +
  xlim(c(24,45))
panelb <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("PKLE"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = SUBPOP)) +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_manual(values = rev(color_code)) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "none", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX2", subtitle = "Day length change on date of flowering") +
  xlim(c(24,45))
panelc <- dyln_latbin2 %>%
  left_join(sites, by = "SITE") %>%
  filter((SITE %in% c("TMPL"))) %>%
  mutate(n = case_when(FL_B4_SOL == "Lengthening" ~ -count,
                       TRUE ~ count)) %>%
  ggplot(aes(x = latbin, y = n, width = 1)) + 
  geom_col(aes(fill = SUBPOP)) +
  coord_flip() +
  facet_wrap(~FL_B4_SOL, ncol = 2, scales = "free_x") +
  scale_fill_manual(values = rev(color_code)) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45),
        legend.position = "right", panel.spacing.x = unit(x = 0, unit = "mm")) +
  scale_y_continuous(expand = c(0.002, 0.002), 
    labels = function(x) signif(abs(x), 3)) +
  ylab("Number of genotypes") +
  xlab("Latitude of origin") + labs(title = "TX3", subtitle = "Day length change on date of flowering") +
  xlim(c(24,45))

dyln_lat_plot <- plot_grid(panela, panelb, panelc, nrow = 1, rel_widths = c(1, 1, 1.318))
save_plot(filename = "../manuscript/plots/Daylength_changes_flowering_date_by_latitude_of_origin_and_subpop.png", plot = dyln_lat_plot, base_asp = 2.4, base_height = 4)
```
